Efficient ancestry and mutation simulation with msprime 1.0

### 2021-09-20 paper.tex
typos

\begin{figure}
\begin{center}
\includegraphics{figures/dtwf-perf}
\end{center}
\caption{\label{fig-dtwf-perf} Comparison of Discrete Time Wright-Fisher
(DTWF) simulation performance in \msprime\ and \ARGON\ (Intel Xeon E5-2680 CPU).
We simulated ancestry for a sample of 1000 haploids from a population of 10000,
and report the (A) total CPU time and (B) maximum memory usage for varying
sequence lengths, and a per-base recombination rate of $10^{-8}$. Each
point is the average over $5$ replicate simulations.
We show observations for \ARGON, \msprime's DTWF implementation (``DWTF'')
and a hybrid simulation of 100 generations of the DTWF followed by
the standard coalescent with recombination model (``DTWF + Hudson'').
Memory usage for \msprime's DTWF and hybrid simulations are very similar.
We ran \ARGON\ with a mutation rate of $0$ and with minimum output options,
to ensure we are measuring only ancestry simulation time.}
\end{figure}

@@

An important goal of \msprime's development model is to maximise
accessibility for prospective users and contributors, and
to encourage diversity in our community.
Gender inequality is a major problem across
the sciences~\citep{shannon2019gender}
and in open source software development~\citep{trinkenreich2021women}.
Within our field, the contribution of women to early computational methods
in population genetics was marginalised~\citep{dung2019illuminating},
and women continue to be under-represented
in computational biology~\citep{bonham2017women}.
The authorship of our paper reflects these trends,
with a skew towards men and affiliations in the USA
and Europe. We know the importance of creating and strengthening networks to
develop and maintain a diverse community of contributors, and we are committed
to fostering a supportive and collaborative environment that helps to address these
inequalities in our field.

@@

The distribution of ancestral alleles
is used to choose the allele present at the root of the tree at each mutated site,
i.e., the \texttt{root\_distribution}.
Mutation models with a finite possible set of alleles have a natural choice
for this distribution---the \emph{stationary distribution} of the mutation process.
(All mutation models are Markovian,
so this may be found as the top left eigenvector of the mutation matrix.)
This is the default in most models,
except, e.g., the \texttt{BinaryMutationModel}, whose alleles are 0 and 1
and always labels the ancestral allele ``0''.
However, mutational processes are not in general stationary,
so we often allow different root distribution to be specified.

>>>>

\begin{figure}
\begin{center}
\includegraphics{figures/dtwf-perf}
\end{center}
\caption{\label{fig-dtwf-perf} Comparison of Discrete Time Wright-Fisher
(DTWF) simulation performance in \msprime\ and \ARGON\ (Intel Xeon E5-2680 CPU).
We simulated ancestry for a sample of 1000 haploids from a population of 10000,
and report the (A) total CPU time and (B) maximum memory usage for varying
sequence lengths, and a per-base recombination rate of $10^{-8}$. Each
point is the average over $5$ replicate simulations.
We show observations for \ARGON, \msprime's DTWF implementation (``DTWF'')
and a hybrid simulation of 100 generations of the DTWF followed by
the standard coalescent with recombination model (``DTWF + Hudson'').
Memory usage for \msprime's DTWF and hybrid simulations are very similar.
We ran \ARGON\ with a mutation rate of $0$ and with minimum output options,
to ensure we are measuring only ancestry simulation time.}
\end{figure}

@@

@@

The distribution of ancestral alleles
is used to choose the allele present at the root of the tree at each mutated site,
i.e., the \texttt{root\_distribution}.
Mutation models with a finite possible set of alleles have a natural choice
for this distribution---the \emph{stationary distribution} of the mutation process.
(All mutation models are Markovian,
so this may be found as the top left eigenvector of the mutation matrix.)
This is the default in most models,
except, e.g., the \texttt{BinaryMutationModel}, whose alleles are 0 and 1
and always labels the ancestral allele ``0''.
However, mutational processes are not in general stationary,
so we often allow a different root distribution to be specified.
